---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**Vignette in development**

In this vignette, we will walk you through the basic usage using the readthis R package for fast reading of output files from various programs, including Mutect2, Strelka, ASCAT, and FACETS. We will use the Strelka VCF files that are installed with the readthis package.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(readthis)

strelka_vcf <- system.file("extdata", "Strelka", "S1.somatic.snvs.vcf.gz", package = "readthis")
```


VCF files with Strelka somatic SNVs can be read with `read_strelka_somatic_snvs()` function. The main argument taken by function is the path, which can be either a path to a single VCF file, vector of paths to many files, or a single path to a directory containing many files.

## Reading a single file

In the simpliest case `path` points a single VCF file:

```{r}
read_strelka_somatic_snvs(strelka_vcf, verbose = FALSE)
```

## Reading list of files

`path` can be a vector of paths to many VCF files

```{r}
strelka_vcf2 <- system.file("extdata", "Strelka", "S2.somatic.snvs.vcf.gz", package = "readthis")
files <- c(S1 = strelka_vcf, S2 = strelka_vcf2)
read_strelka_somatic_snvs(files, verbose = FALSE)
```

## Reading all files from the directory

```{r}
strelka_dir <- system.file("extdata", "Strelka", package = "readthis")
read_strelka_somatic_snvs(strelka_dir, verbose = FALSE)
```

`readthis` contains methods for bulk reading of output files from some other programs. To see the list of  functions implemented in the package go to [Reference](https://pawelqs.github.io/readthis/reference/index.html) page.
